### 2019-03-06 micro4_1.tex
Micro

	\item No, they cannot be. Firstly, obviously that the greatest amount of efforts the workers can spend on education at any pooling equilibria cannot exceed $\bar{e}$ (see Fig. \ref{fig5}) because for any other education level $e > \bar{e}$ and respective equilibrium wage level $w(e) = \lambda \theta_H + (1 - \lambda)\theta_L + e$ the low type worker prefers $e = 0$.

@@

Suppose now that the pooling equilibrium is attained at education level $e' < \bar{e}$ (see Fig. \ref{fig5}). The corresponding indifference curves and wage schedule is depicted on the Fig. \ref{fig5}. Both types are maximizing their utilities, given the wage schedule and equilibrium wages are consistent with beliefs. And this pooling equilibrium Pareto dominates the one with $e = \bar{e}$ (and also each equilibrium with $e' < e < \bar{e}$) because firms have the same zero expected profit, but both types attain higher indifference curves and consequently higher utility levels at $e'$. However, there always exists an education level $\underline{e} = \theta_H$ such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig. \ref{fig5}). And each pooling  equilibrium level of education $e'' < \underline{e}$ will yield the high type worker strictly less utility. Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{e}$ corresponds to lower indifference curve for high type. Similarly for low type workers there exists an education level $\underline{\underline{e}} = \theta_L$ such that indifference curve of low type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig. \ref{fig5}). And each pooling  equilibrium level of education $e''' < \underline{\underline{e}}$ will yield the low type worker strictly less utility. Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{\underline{e}}$ corresponds to lower indifference curve for low type. As a result, there is a full Pareto order for pooling equilibrium for $e \in [\theta_H, \bar{e}]$ and for $e \in [0, \theta_L]$, for $e \in (\theta_L, \theta_H)$ there is no Pareto order, since some pooling equilibria are better for low type, but are worse for high type and vice-versa.
\item Highest pooling equilibrium payoff for the high type is delivered at pooling equilibrium education level $e^* = \theta_H$

@@

>>>>

	\item No, they cannot be. Firstly, obviously that the greatest amount of efforts the workers can spend on education at any pooling equilibria cannot exceed $\bar{e}$ (see Fig. \ref{fig5}) because for any other education level $e > \bar{e}$ and respective equilibrium wage level $w(e) = \lambda \theta_H + (1 - \lambda)\theta_L + e$ the low type worker prefers $e = 0$. (here $\bar{e} = \theta_L\left(1 + \sqrt{1 + \frac{2\lambda(\theta_H - \theta_L)}{\theta_L}}\right) > \theta_L$)

@@

Suppose now that the pooling equilibrium is attained at education level $e' < \bar{e}$ (see Fig. \ref{fig5}). The corresponding indifference curves and wage schedule is depicted on the Fig. \ref{fig5}. Both types are maximizing their utilities, given the wage schedule and equilibrium wages are consistent with beliefs. And this pooling equilibrium Pareto dominates the one with $e = \bar{e}$ (and also each equilibrium with $e' < e < \bar{e}$) because firms have the same zero expected profit, but both types attain higher indifference curves and consequently higher utility levels at $e'$. However, there always exists an education level $\underline{e} = \theta_H$ (which is wlog $< \bar{e}$) such that indifference curve of high type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig. \ref{fig5}). And each pooling  equilibrium level of education $e'' < \underline{e}$ will yield the high type worker strictly less utility. Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{e}$ corresponds to lower indifference curve for high type. Similarly for low type workers there exists an education level $\underline{\underline{e}} = \theta_L$ such that indifference curve of low type is tangent to the line $\lambda \theta_H + (1 - \lambda)\theta_L + e$ at this point (see Fig. \ref{fig5}). And each pooling  equilibrium level of education $e''' < \underline{\underline{e}}$ will yield the low type worker strictly less utility. Since each point at line $\lambda \theta_H + (1 - \lambda) \theta_L + e$ for $e < \underline{\underline{e}}$ corresponds to lower indifference curve for low type. As a result, there is a full Pareto order for pooling equilibrium for $e \in [\theta_H, \bar{e}]$ and for $e \in [0, \theta_L]$, for $e \in (\theta_L, \theta_H)$ there is no Pareto order, since some pooling equilibria are better for low type, but are worse for high type and vice-versa.

Case $\theta_H > \bar{e}$ is analogous, for $e \in (\theta_L, \theta_H)$ equilibria cannot be Pareto ranked.
\item Dominance requirement requires firms to prescribe positive $\mu(e)$ only to such $e$ that is not a strictly dominated action for high type. In the sense that there is no action $e'$ which yields the high type strictly greater payoff regardless of what equilibrium wages are. Assumption that $\Omega$ in non-empty means that firstly $\theta_H < \bar{e}$ and on the top of that the indifference curve of high type, which goes through $(\bar{e}, \theta_H + \bar{e})$ lies below the indifference curve which is tangent to the line $\lambda \theta_H + (1 - \lambda) \theta_L$ at the point $\theta_H$ (see Fig. \ref{fig6}) because otherwise, the high type chooses the level $\max\ \{\theta_H, \bar{e}\}$ and low type chooses $0$ and $\Omega$, the set of pooling equilibria, is empty.
	\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{plotdraft}
	\caption{}\label{fig6}
\end{figure} 
The highest pooling equilibrium payoff for the high type is delivered at the education level $e = \theta_H$. In our case this equilibrium belongs to $\Omega$ because neither of types have $e = \theta_H$ as a strictly dominated action. On the Fig. \ref{fig6} there is a wage schedule for which $e_H = e_L = \theta_H$ is the best response for both types (i.e. there is no action which strictly dominates this action).
	\end{enumerate}
\end{enumerate}
\section*{4\ Exams}
Karen would like her friend Tim to pass his exams. Tim may choose to work at a
private cost of $c$. If he slacks, he fails. If he works, he passes with probability $\theta$. Hence $\theta$
is Tim's $ability$. If he passes, Karen receives a private benefit of $V > 0$ and Tim receives
a private benefit of $U > 0$.
Karen offers to pay a reward $b \in [0; V]$ to Tim if and only if he passes. Hence Karen
receives a payoff $V-b$ if Tim passes, and zero otherwise. Tim receives a payoff of zero if
he slacks, $-c$ if he works and fails, and $U + b-c$ if he works and passes. For simplicity,
assume that Tim will choose to work whenever he is otherwise indifferent between working
and slacking.
Tim's ability is initially unknown; $\theta_H$ with probability $\alpha$ and $\theta_L$ otherwise where
\begin{align*}
1 > \theta_H > \theta_L > 0, [\alpha \theta_H + (1-\alpha)\theta_L] U < c < \theta_L(U + V ) \text{ and } c < \theta_H U
\end{align*}
We refer to Tim's beliefs about $\theta$ as his "confidence". Begin by assuming that neither
player observes $\theta$. Hence Karen begins by choosing a bonus $b$. Tim observes this bonus,
and then chooses whether to work
\begin{enumerate}
	\item What bonus will Karen choose to offer?

	Karen now observes Tim's ability $\theta$, but Tim does not. Hence, Nature begins by
	choosing $\theta \in \{\theta_H; \theta_L\}$. Next, Karen observes $\theta$ and offers a bonus $b_H$ (if $\theta = \theta_H$) or $b_L$
	(if $\theta = \theta_L$). Finally, Tim observes $b$, but not $\theta$, and then chooses whether to work. We seek pure strategy perfect Bayesian Nash equilibria.
	\item Show that there is no separating equilibrium, and hence $b^*_H = b^*_L$ in any equilibrium.
	\item Does the standard sorting (single crossing) condition hold in part 2?

	Suppose now that Karen offers Tim an unconditional gift $w \ge 0$ as well as the bonus
	$b \ge 0$. Hence she pays $w+b$ if Tim passes, and $w$ if he fails. So Karen offers $(w_H; b_H)$ when
	she observes $\theta_H$ and $(w_L; b_L)$ when she observes $\theta_L$. We seek a separating equilibrium.

	\item Show that $w^*_L = 0$ and $b^*_L = \frac{c}{\theta_L} - U$ in any separating equilibrium. Show that
	together with $w^*_H = c - \theta_LU$ and $b_H^* = 0$ this constitutes an equilibrium.
	\item Briefly interpret this equilibrium
\end{enumerate}
	\item Since Karen gets 0 if Tim slacks and if he works she in the worst case gets 0, she should offer such a $b$ that Tim would be indifferent between work and slack, and as a result chooses to work.
	\alpha(\theta_H(U + b - c) &- (1 - \theta_H)c) + (1 - \alpha)(\theta_L(U + b - c) - (1 - \theta_L)c) = 0\\
	&(U+b)(\alpha\theta_H + (1 - \alpha)\theta_L) = c\\
	&b = \frac{c}{\alpha\theta_H + (1 - \alpha)\theta_L} - U
	\item In separating equilibrium Tim can perfectly distinguish whether $\theta$ is high or low, hence in case of high $\theta$ he will work only if:
	(U + b_H)\theta_H - c \ge 0\\
	b_H \ge \frac{c}{\theta_H} - U < 0
	i.e. if $\theta = \theta_H$ Karen's optimal offer is $b_H = 0$. If $\theta = \theta_L$ then Tim will work only if
	b_L \ge \frac{c}{\theta_L} - U
	and hence optimal bonus which Karen offers is:
	b_L = \max\ \left\{\frac{c}{\theta_l} - U, 0\right\}
	but if $b_L = 0$ then equilibrium becomes pooling and Tim cannot distinguish between states. If $\frac{c}{\theta_l} - U > 0$ then Karen will want to deviate and offer $b_L = 0$ in order to convince Tim that he is of high type. Hence at each equilibrium $b^*_H = b^*_L$.
	\item  As always at separating equilibrium Tim can perfectly distinguish between states. That means that at low state he will work only if
	\theta_L(U + b_L) + w_L - c \ge w_L
	hence Karen should offer such $(b_L, w_L)$ that:
	\theta_L b_L = c - \theta_LU
	In high state it is profitable for Tim to accept any positive offer $(b_H, w_H)$ and it would be optimal for Karen to offer $(0, 0)$ but it cannot be a separating equilibrium because at low state Karen will have a profitable deviation, namely $(0, 0)$ and equilibrium becomes pooling. So, at the separating equilibrium $(w_H, b_H)$ should be such that Karen does not want to deviate at low state. That is:
	\theta_L(V - b_H - w_H) + (1 - \theta_L)(-w_H) &\le \theta_L(V - b_L - w_L) + (1 - \theta_L)(-w_L)\\
	\theta_L V - \theta_Lb_H - w_H &\le \theta_L(V + U) - c - w_L
	hence
	\begin{align}\label{eq1}
	\theta_L b_H + w_H = c +w_L - \theta_L U
	\end{align}
	and finally Karen should choose such $(b_H, w_H), (b_L, w_L)$ which maximize her expected payoff. At low state she should maximize:
	\theta_L (V - b_L) - w_L = \theta_L V - c + \theta_LU - w_L \to \underset{w_L}{\max}
	obviously the solution is $w_L = 0$. Similarly, using \eqref{eq1}, at high state she should maximize:
	\theta_HV - \theta_H b_H - w_H = \theta_HV - \theta_Hb_H - c + \theta_LU + \theta_L b_H \to \underset{b_H}{\max}\\
	\theta_H V + \theta_l U - c - (\theta_H - \theta_L)b_H \to \underset{b_H}{\max}
	the solution is $b_H = 0$. Thus, the equilibrium is
	w^*_L = 0,\ b^*_L = \frac{c}{\theta_L} - U\\
	w^*_H = c - \theta_LU,\ b^*_H = 0
	as we sought.
	\item The interpretation is straightforward. Without unconditional gift Karen pays her bonus ex post, as a result she had incentives to deviate and underpay Tim, to cheat him that he is of high type. Whereas the introduction of the gift which Karen pays ex ante can be viewed as a commitment that Karen will not cheat Tim and will not make his expected payoff negative when Tim is of low type.

@@

\section*{5\ Beliefs}
The definition of the world $arbitrary$ is the following:
\text{based on random choice or personal whim, rather than any reason or system}
\end{align*}First of all such beliefs are not fully arbitrary, they should be consistent with strategies, players play. That is, given those beliefs and actions the player should undertake according to them, players do not have profitable unilateral deviations from the equilibrium path. But even if we actually can arbitrary (to the aforementioned extent) choose beliefs for the off equilibrium path info sets, conclusions are not arbitrary in the slightest. They are fully consistent with this particular concept of equilibrium. We do not conclude that tomorrow will be rain because some information set at the game we study is reached with zero probability at the equilibrium. It is the example of arbitrary conclusion. But those conclusions we usually draw are correct for the model and concept of equilibrium we chose. If we do not like these conclusions we can only choose another concept of equilibrium, use some refinements for example, but we cannot say that conclusions are arbitrary, they are valid, correct for this concept and not random.
\section*{6\ Wage Inequality}
